---
title: "SWOT_river_US_graphic"
author: "Merritt Harlan"
date: "2023-04-21"
output: html_document
---

## Set up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

Define libraries here. 

```{r libraries, warning=FALSE, message=FALSE}
# Load libraries
library(sf)
library(dplyr)
library(ggplot2)
library(downloader)
library(dataRetrieval)
library(ggthemes)
library(scales)

#library(tidyverse)
#library(readr) # best for reading and writing csvs
#library(downloader)
#
#library(raster)
#library(sf)
#library(dplyr)

# These are used for the layout of the viz
#library(ggpubr)
library(cowplot) # for laying out the final plot
#library(sysfonts) # for text editing
#library(showtext) # for adding in google fonts
#library(magick) # for adding logo
```



## Load files

```{r load data}
# load US outline
us_outline_zip <- download("https://www2.census.gov/geo/tiger/GENZ2022/shp/cb_2022_us_nation_20m.zip", destfile = "in/us_outline.zip")

unzip("in/us_outline.zip", exdir = "in/us_outline")

us_outline <- read_sf("in/us_outline/cb_2022_us_nation_20m.shp") %>% st_transform(4326)
ak_outline <- spData::alaska %>% st_transform(4326)
cont_outline <- spData::us_states %>% st_transform(4326)

#load SWORD database, filter to US
SWORD_US <- read_sf("../SWORD/SWORD_v15_gpkg/gpkg/na_sword_reaches_v15.gpkg") %>% 
  st_intersection(us_outline) %>% 
  dplyr::select(reach_id, reach_len, wse, width, swot_obs)

SWORD_AK <- SWORD_US %>% st_intersection(ak_outline) %>% dplyr::select(reach_id, reach_len, wse, width, swot_obs)
SWORD_cont <- SWORD_US %>% st_crop(st_bbox(cont_outline))

#load network NHD flowline in continental US


#load current gage data
USGS_gage_data <- data.frame()
for(hc in 1:22){
  huc_id = sprintf("%02d", as.numeric(hc))
  huc_gage = whatNWISsites(parameterCd = "00060", huc = huc_id, period= "P52W")
  USGS_gage_data = rbind(USGS_gage_data, huc_gage)
}

USGS_gage_data_sf <- st_as_sf(USGS_gage_data, coords = c("dec_long_va", "dec_lat_va"), crs = 4269) %>% 
  dplyr::select(site_no)

USGS_gage_data_AK <- whatNWISsites(stateCd = "AK", period= "P52W", parameterCd = "00060")
USGS_gage_data_AK_sf <- st_as_sf(USGS_gage_data_AK, 
                                 coords = c("dec_long_va", "dec_lat_va"), 
                                 crs = 4269) %>% dplyr::select(site_no)
nrow(USGS_gage_data)

# load PLD database, filter to US
# PLD_AR_81 <- read_sf("../PLD/PLD_AR_81.gpkg") %>% st_intersection(us_outline)
# PLD_AR_82 <- read_sf("../PLD/PLD_AR_82.gpkg") %>% st_intersection(us_outline)
# PLD_NA_71 <- read_sf("../PLD/PLD_NA_71.gpkg") %>% st_intersection(us_outline)
# PLD_NA_72 <- read_sf("../PLD/PLD_NA_72.gpkg") %>% st_intersection(us_outline)
# PLD_NA_73 <- read_sf("../PLD/PLD_NA_73.gpkg") %>% st_intersection(us_outline)
# PLD_NA_74 <- read_sf("../PLD/PLD_NA_74.gpkg") %>% st_intersection(us_outline)
# PLD_NA_75 <- read_sf("../PLD/PLD_NA_75.gpkg") %>% st_intersection(us_outline)
# PLD_NA_76 <- read_sf("../PLD/PLD_NA_76.gpkg") %>% st_intersection(us_outline)
# PLD_NA_77 <- read_sf("../PLD/PLD_NA_77.gpkg") %>% st_intersection(us_outline)
# PLD_NA_78 <- read_sf("../PLD/PLD_NA_78.gpkg") %>% st_intersection(us_outline)

```

## Set up main plot

```{r plotting}
# Load some custom fonts and set some custom settings
sysfonts::font_add_google("Source Sans Pro")
supporting_font <- "Source Sans Pro"
showtext::showtext_opts(dpi = 300, regular.wt = 200, bold.wt = 700)
showtext::showtext_auto(enable = TRUE)

# Define colors
background_color = "white"
font_color = "dodgerblue4"

# The background canvas for your viz
canvas <- grid::rectGrob(
  x = 0, y = 0, 
  width = 16, height = 9,
  gp = grid::gpar(fill = background_color, alpha = 1, col = background_color)
)

# Load in USGS logo (also a black logo available)
usgs_logo <- magick::image_read("../../Research/Chart_challenge/chart-challenge-23/usgs_logo_black.png")

# Plots
sword_map <- ggplot() +  
  geom_sf(data=us_outline %>% st_crop(st_bbox(SWORD_US)), color = "grey") + 
  geom_sf(data=SWORD_US %>% filter(width > 100), color = "dodgerblue4", linewidth = 1.25) +
  geom_sf(data=SWORD_US %>% filter(width < 100), color = "dodgerblue4", linewidth = 0.75) +
  geom_sf(data=USGS_gage_data_sf %>% st_transform(4326) %>% st_crop(st_bbox(SWORD_US)), color = "dodgerblue", size = 0.5, alpha = 0.3) +
  theme_map()
  #theme(legend.position="right")


```


## Produce final plot

Here, use `cowplot` and `ggsave` to create the final viz for sharing out on Twitter. This template includes adding the USGS logo, title, text, etc.

```{r cowplot, fig.width = 16, fig.height = 9}
#Static Image
ggdraw(ylim = c(0,1), # 0-1 scale makes it easy to place viz items on canvas
       xlim = c(0,1)) +
  # a background
  draw_grob(canvas,
            x = 0, y = 1,
            height = 9, width = 16,
            hjust = 0, vjust = 1) +
  # satellite image
  draw_plot(sword_map,
            x = 0.00,
            y = 0.00,
            width = 0.98) +
  # explainer text
  draw_label(paste('Merritt Harlan, USGS',
                   'NHD: National Hydrography Dataset',
                   'SWOT: Surface Water Ocean Topography satellite- launched December 16, 2022', 
                   sep = '\n'),
             fontfamily = supporting_font,
             x = 0.96,
             y = 0.03,
             size = 10,
             hjust = 1,
             vjust = 0,
             color = "black")+
  # Title
  draw_label(paste('Satellite Monitoring of Rivers:'),
             x = 0.40,
             y = 0.92,
             hjust = 0,
             vjust = 0,
             lineheight = 0.75,
             fontfamily = supporting_font,
             fontface = 'bold',
             color = font_color,
             size = 40) +
 # Subtitle
  draw_label(paste('Surface Water Ocean Topography (SWOT) Satellite'),
             x = 0.40,
             y = 0.87,
             hjust = 0,
             vjust = 0,
             lineheight = 0.75,
             fontfamily = supporting_font,
             color = font_color,
             size = 24) +
   # Annotation
  draw_label(paste('SWOT Data Products:'),
             x = 0.05,
             y = 0.55,
             hjust = 0,
             vjust = 0,
             lineheight = 0.75,
             fontfamily = supporting_font,
             fontface = "italic",
             color = "dodgerblue3",
             size = 20) +
  # Annotation
  draw_label(paste('Water surface extent',
                   'Water surface elevation',
                   'Streamflow*', sep = '\n'),
             x = 0.05,
             y = 0.47,
             hjust = 0,
             vjust = 0,
             lineheight = 0.75,
             fontfamily = supporting_font,
             color = "dodgerblue3",
             size = 16) +
  # Annotation
  draw_label(paste('Rivers in Numbers:'),
             x = 0.40,
             y = 0.78,
             hjust = 0,
             vjust = 0,
             lineheight = 0.75,
             fontfamily = supporting_font,
             fontface = "italic",
             color = "dodgerblue3",
             size = 20) +
  # Annotation
  draw_label(paste('NHD flowlines:          1,000,000   (%)      100,000 mi   (%)',
                   'USGS streamflow gages:          10,000   0.001      ',
                   'SWOT river reaches:          16,000   0.0016    42.3',
                   'SWOT river reaches > 100 m*:          8,500   0.00085   36.1', sep = '\n'),
             x = 0.70,
             y = 0.68,
             hjust = 1,
             vjust = 0,
             lineheight = 0.75,
             fontfamily = supporting_font,
             color = "dodgerblue3",
             size = 16) +
  # Add logo
  draw_image(usgs_logo,
             x = 0.96,
             y = 0.10,
             width = 0.10,
             hjust = 1, vjust = 0,
             halign = 0, valign = 0)

# Save the final image in Twitter's 16 by 9 format
# !! Use format for saving with the date of your prompt: 
#         YYYYMMDD_prompt_name ()
# e.g. `20230101_part-to-whole-cnell.png`
ggsave(filename = "out/SWOT_US_Rivers.png", 
       width = 14, height = 9, dpi = 300)
```

## Final GIF

```{r gif}
##### Create 3D GIF for output #####

# Set up parameters for loop to create GIF
n_frames <- length(dswe_images)
img_frames <- paste0("out/cache/dswe_uncertainty_", seq_len(n_frames), ".png")


for(i in 1:n_frames){
  dswe_index = arrange(tanana_dswe_cleaned, width_ft)[i,]
  date = dswe_index$Date
  flow_index = tanana_flow_cleaned %>% filter(Date == date)
  
  
  filtered_image = lapply(dswe_images, function(x) filter(x, Date == date))
  dswe_image = data.frame()
  for(x in 1:length(filtered_image)){
    if(nrow(filtered_image[[x]])>0){
      dswe_image= filtered_image[[x]]
    }
  }
  
  dswe_image$DSWE = factor(dswe_image$DSWE, levels = c("", "water - high confidence", "water - moderate conference", "potential wetland", "water/wetland - low confidence"))
  
  dswe_plot <- ggplot() +  
    geom_tile(data=dswe_image, aes(x=x, y=y, fill= DSWE), alpha=0.8) + 
    coord_equal() +
    scale_fill_manual(name = "Classification", values = c("white", "dodgerblue4", "dodgerblue", "lightskyblue", "lightblue1")) +
    theme_map() +
    theme(legend.position="right")
  
  width_plot_new <- width_plot + geom_point(data = dswe_index, aes(x = wse_ft, y = width_ft), 
                                            size = 7, shape = 19, color = "dodgerblue")

  ggdraw(ylim = c(0,1), # 0-1 scale makes it easy to place viz items on canvas
       xlim = c(0,1)) +
  # a background
  draw_grob(canvas,
            x = 0, y = 1,
            height = 9, width = 16,
            hjust = 0, vjust = 1) +
  # satellite image
  draw_plot(dswe_plot,
            x = 0.00,
            y = 0.00,
            width = 0.98) +
  # width height plot
  draw_plot(width_plot_new,
            x = 0.03,
            y = 0.01,
            height = 0.30,
            width = 0.20) +
  # height flow plot
  draw_plot(altim_plot,
            x = 0.24,
            y = 0.01,
            height = 0.30,
            width = 0.20) +
  # height flow plot
  draw_plot(flow_plot,
            x = 0.45,
            y = 0.01,
            height = 0.30,
            width = 0.30) +
  # Alaska inset map
  draw_plot(AK_inset,
            x = 0.75,
            y = 0.72,
            height = 0.25,
            width = 0.25) +
  # explainer text
  draw_label(paste('Merritt Harlan, USGS', 
                   'Data: Smith, T. L., et al., 2022, Computed streamflow using',
                   'satellite data for the Copper, Nushagak, Tanana, Yukon,', 
                   'Susitna, and Knik, Koyukuk Rivers, Alaska, 2008-2021:',
                   'USGS data release, https://doi.org/10.5066/P94LLG4R.',
                   'Tanana River in Fairbanks, AK- USGS Site 15485500', sep = '\n'),
             fontfamily = supporting_font,
             x = 0.96,
             y = 0.03,
             size = 10,
             hjust = 1,
             vjust = 0,
             color = "black")+
 # Title
  draw_label(stringr::str_wrap("Estimating streamflow from satellites",60),
             x = 0.03,
             y = 0.92,
             hjust = 0,
             vjust = 0,
             lineheight = 0.75,
             fontfamily = supporting_font,
             fontface = 'bold',
             color = font_color,
             size = 40) +
  # Annotation
  draw_label(paste("By relating optical satellite imagery, which can give us an estimate of river width,",
                   "to satellite altimetry, which gives us an estimate of river elevation, we can approximate",
                   "streamflow in large rivers. Although this data can have high uncertainty, combined with ",
                   "knowledge of rivers and ground data, satellites can help us improve monitoring.", sep = '\n'),
             x = 0.05,
             y = 0.77,
             hjust = 0,
             vjust = 0,
             lineheight = 0.75,
             fontfamily = supporting_font,
             color = font_color,
             size = 18) +
  # Add logo
  draw_image(usgs_logo,
             x = 0.96,
             y = 0.15,
             width = 0.10,
             hjust = 1, vjust = 0,
             halign = 0, valign = 0)

  # Save the final image in Twitter's 16 by 9 format
  # !! Use format for saving with the date of your prompt: 
  #         YYYYMMDD_prompt_name ()
  # e.g. `20230101_part-to-whole-cnell.png`
  ggsave(filename = img_frames[i], 
         width = 14, height = 9, dpi = 300)
}

# Build GIF using cached images - add labels here to control size better
magick::image_read(img_frames) %>%
  magick::image_write_gif(path="out/20230429_monochrome_mharlan.gif",delay=12/n_frames) # Number here is duration

```

## Supporting information

### Key takeaways of this viz (1-2 sentences each)

1. Satellite optical and altimetry data can be combined to estimate streamflow, as shown here on the Tanana River in Fairbanks Alaska. We can use the classified water extent, mapped here, to estimate river width by dividing water area and the length of the river, and match this data with overlapping satellite altimeters to build relationships between river width, elevation, and streamflow. 

### Data source(s)

Landsat Collection 2 Level-3 Dynamic Surface Water Extent (DSWE) Science Products courtesy of the U.S. Geological Survey.

Jones, J.W., 2019. Improved Automated Detection of Subpixel-Scale Inundation—Revised Dynamic Surface Water Extent (DSWE) Partial Surface Water Tests. Remote Sens., 11, 374 https://doi.org/10.3390/rs11040374.

Smith, T. L., et al., 2022, Computed streamflow using satellite data for the Copper, Nushagak, Tanana, Yukon, Susitna, and Knik, Koyukuk Rivers, Alaska, 2008-2021: United States Geological Survey data release, https://doi.org/10.5066/P94LLG4R.

Tanana River in Fairbanks, AK- USGS Site 15485500

